---
title: "dataset36_withmed"
author: "Fay Yan"
date: "2024-07-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

load data 
dataset: newDatabase_36_ANON

```{r, warning=FALSE}
library(readxl)
newDatabase_36_ANON <- read_excel("~/Desktop/newDatabase_36_ANON.xlsx", 
    sheet = "Asymptomatic_CAC", skip = 2)
```

select column: (including variables for medication)

```{r}
library(dplyr)
ANON36_selected <- newDatabase_36_ANON[, c(1:6,11,40,41,43,44,45,47,49,
                                           50:57,61:64,65:66,96,107:115)]
```

For risk factors, encode to binary (change NA to 0)

```{r}
ANON36_selected <- ANON36_selected %>%
  mutate(across(15:22, ~ ifelse(is.na(.) | . == "", 0, .))) 
```

add column Male:

```{r}
ANON36_selected <- ANON36_selected %>%
  mutate(Male = ifelse(Gender == "Male", 1, 0))
```

handle Race column:
"American Indian or Alaskan Native" = 1
"Asian" = 2
"Black" = 3
"Hispanic" = 4
"Other" = 5
"Unknown" = 6
"White" = 7
NA = missing 

```{r}
# Step 1: Convert "NA" strings to actual NA values
ANON36_selected$Race[ANON36_selected$Race == "NA"] <- NA
# Step 2: Convert Race to a factor
ANON36_selected$Race <- as.factor(ANON36_selected$Race)
# Step 3: Manually encode Race as numeric, keeping NA as NA
ANON36_selected$Race_encoded <- ifelse(is.na(ANON36_selected$Race), NA, as.numeric(ANON36_selected$Race))
```

For medication columns, encode to binary (change NA to 0)

```{r}
ANON36_selected_withmed <- ANON36_selected %>%
  mutate(across(c(`Beta Blockers`, `Antianginal Agents`, `Antihyperlipidemic`, `Calcium Blockers`, `Antihypertensive`,
                  `Antidiabetic`, `Antiarrhythmic`), 
                ~ ifelse(grepl("prior|on date", ., ignore.case = TRUE), 1, 0)))
```


MACE dataset:
when either MACE or Death = 1, result = 1; otherwise 0

```{r}
ANON36_MACE <- ANON36_selected_withmed[, c(1,4,7:26,29:40)] 
ANON36_MACE <- ANON36_MACE %>%
  mutate(result = if_else(!is.na(`First MACE or Death (1=MACE, 2=Death)`), 1, 0))
ANON36_MACE_output <- ANON36_MACE[, -c(23)]
```


```{r}
write.csv(ANON36_MACE_output, "/Users/Student/Desktop/ANON36_MACE_output.csv", row.names = FALSE)
```

HF dataset:
when CHF date after exam date, result = 1; otherwise 0

```{r}
ANON36_HF <- ANON36_selected_withmed[, c(1:2,4,7:28,30:40)]
ANON36_HF$`Exam Completion Timestamp` <- as.Date(ANON36_HF$`Exam Completion Timestamp`, format = "%Y-%m-%d")
ANON36_HF$`CHF Dx Date` <- as.Date(ANON36_HF$`CHF Dx Date`,format = "%Y-%m-%d")
ANON36_HF <- ANON36_HF %>%
  mutate(is_later = case_when(
    !is.na(`Exam Completion Timestamp`) & !is.na(`CHF Dx Date`) ~ `CHF Dx Date` > `Exam Completion Timestamp`,
    TRUE ~ NA
  ))
ANON36_HF <- ANON36_HF %>%
  mutate(result = if_else(is.na(is_later) | is_later == FALSE, 0, 1))
ANON36_HF_output <- ANON36_HF[, -c(2,24,25,37)]
```

```{r}
write.csv(ANON36_HF_output, "/Users/Student/Desktop/ANON36_HF_output.csv", row.names = FALSE)
```

death dataset:
when Death = 1, result = 1; otherwise 0

```{r}
ANON36_death <- ANON36_selected_withmed[, -c(2,3,5,6,27:28)]
ANON36_death <- ANON36_death %>%
  mutate(`First MACE or Death (1=MACE, 2=Death)` = ifelse(is.na(`First MACE or Death (1=MACE, 2=Death)`), 0, `First MACE or Death (1=MACE, 2=Death)`)) %>%
  mutate(result = ifelse(`First MACE or Death (1=MACE, 2=Death)` == 2, 1, 0))
ANON36_death_output <- ANON36_death[, -c(23)]
```


```{r}
write.csv(ANON36_death_output, "/Users/Student/Desktop/ANON36_death_output.csv", row.names = FALSE)
```
